---
title: "Best States: Black, Hispanic, Female Tech -- V0.9"
author: "Roy L Beasley, PhD -- DLL Editor"

output: 
 prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
    toc_depth: 4
    fig_width: 7
    fig_height: 6
          
---

```{r, echo=FALSE, cache=FALSE}
    
    load("dfRaceResidenceAndShares.rda")
    load("dfCensus2.rda") 
    load("dfTab1A1B2A2B3.rda")
    load("dfTab4.rda")   
    load("dfMap4.rda")
    load("dfPlot5Tab5beta1000.rda")
    load("dfTab67A7B7C7D7E7F8A8B8C8D8E8F.rda")
    
    library(prettydoc)
    library(maps)    
    library(mapproj) ### needed by ggplot2, but not installed automaticallly
    library(ggplot2)
    library(gridExtra)
    ### library(printr)
    library(grid)
    library(gtable)

    ### Intending to use Yihui Xie's printr package for better tables later
    ### install.packages("devtools")
    ### library(devtools)
    ### install_github("yihui/printr")
    ### library(printr)
    ### detach('package:printr', unload = TRUE)
    
tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = blues9[1:7], col=NA),
            fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="navyblue", fontface=4L)),
  rowhead=list(fg_params=list(col="orange", fontface=3L)))
```
## Overview   
My principal findings were straight-forward:   

* First, in 2014 there were approximately `r prettyNum(dfRaceResidenceAndShares[1,2], big.mark=",", scientific=FALSE)` employees in the entire U.S. information technology sector ==>  `r prettyNum(dfRaceResidenceAndShares[1,4], big.mark=",", scientific=FALSE)` were Black; `r prettyNum(dfRaceResidenceAndShares[1,6], big.mark=",", scientific=FALSE)` were Hispanic; and `r prettyNum(dfTable2B[1,3], big.mark=",", scientific=FALSE)` were female. (See Table 1A, Table 1B, Table 2A, and Table 2B.)      

* Second, the number of tech employees who were Black was higher in states that had larger Black populations and smaller in states that had smaller Black populations. The same relationship between population size and tech employment held true for Hispanics and females. (See Table 4B, Table 4D, and Table 6)

* Third, if all other things were equal, one would expect that the percentage of a state's tech sector that was Black would be about the same as the percentage of the state's total population that was Black. One would expect similar parity for the percentages of Hispanics and for females. But all things were not equal, so I found that the percentage of the tech sector of most states that was Black, Hispanic, or female was substantially smaller than the percentage of the state's total population that was Black, Hispanic, or female. By contrast, I found that the percentage of state tech sectors that was Asian was always greater than their population percentage and ranged as high as `r dfTable6["Asian", "max"]` times their population percentage. (See Tables 4C and 6)  

* Fourth, females techs only held `r prettyNum(dfTable2B[2,3])` percent of the nation's `r prettyNum(dfTable2B[1,1], big.mark=",", scientific=FALSE)` tech positions. After considering the disproportionate number of tech positions allocated to Asian techs, I split the female tech category into Asian females and non-Asian females. I then found that Asian females also occupied disproportionately larger shares of the tech sectors in most states than their share of each state's population. Indeed, their tech shares ranged as high as 6.29 times their shares of state populations. (See Table 4E and Table 6.) By contrast, non-Asian females occupied much smaller shares of the state tech sectors than their shares of the state populations. (See Tables 2B, 3 and 6.)

* Fifth, the strengths of these relationships varied widely: some states offered substantially larger employment opportunities for Blacks, Hispanics, and female techs than others. (See Tables 8B, 8D, 8E, and 8F)   

* Sixth, California and New York, the states that employed the largest and second largest number of information technology professionals, were "best states" for White information technology professionals; but neither was a "best state" for Blacks, Hispanics, Asians, Asian females, or non-Asian females in tech.
<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>   


## Motivation   
My motivation for conducting this study influenced my methods. The recent barrage of <b>[annual reports](https://techcrunch.com/tag/diversity-report/)</b> from some of the nation's leading information technology companies in Silicon Valley said the primary reason why those companies employed so few Black, Hispanic, and female information technology professionals was because there were so few available. And the reason they weren't available was because there were so few Black, Hispanic, and female tech students in the nation's academic pipelines.    

<b>A. Best Practices</b>   
When reading the Silicon Valley reports I found myself shaking my head while wondering, "What numbers did they have to back up these assertions? And where did they get them from?" Whereas I would have said that the pipelines were half-full, these reports implied that the pipelines were perilously close to empty. But if the pipelines were close to empty, there shouldn't be many Black, Hispanic, or female employees in the U.S. tech sector, right? Which led me to wonder how many Black, Hispani, and femalec techs were out there and where were they located. 

<b>But what concerned me the most was the lack of context for assessing Silicon Valley's diversity numbers. Were they as good as could be expected? Were they worse? ... And compared to what?</b> Compared to the rest of California? Compared to other states in other parts of the country? The personal data that I had accumulated from more than 50 years in tech informed me that tech was much more diverse in the Eastern and Southern states. But personal data, while a useful source of productive hypotheses, is not credible, systematic proof. As a clever tech pundit recently proclaimed, <i>"The plural of personal anecdote is not data" ... :-)</i>

One of the most useful lessons that I learned as a consultant to the National Center for Productivity way back in the late 1970s was that the application of performance metrics, even flawed metrics, to just about any significant human activity invariably identified a broad middle that was above a small tier of low performers and below another small tier of high performers. Further inquiry usually disclosed that the high performers were doing things quite differently than the middle and low performers. Happily, the performance of the lower tiers could be improved by their adoption of the high performers' <b>best practices</b> if they made appropriate adaptations of the best practices to their local working conditions. 

Given this insight, I set out to measure the performance of all 50 states and the District of Columbia with regards to the opportunities their tech sectors provided for Black, Hispanic, and female professionals. If Silicon Valley proved to be a low performer in these national contexts, its executives would be unlikely sources of best practices that could improve the Valley's performance.

<b>B. Reproducible Reports</b>   
One of my most important take-aways from completing the <b>[Johns Hopkins "Data Science"](https://www.coursera.org/specializations/jhu-data-science)</b> online certificate program a few months ago was the notion that reports should be "reproducible". Authors should not only describe their methods for collecting and analyzing data; they should provide step-by-step descriptions of their methods (including copies of their code) so that interested readers could readily reproduce the results on their own computers. Of course, the best researchers have always tried to achieve this objective and have often succeeded. What was new for me was the specific tools that the statisticians who taught the Hopkins courses recommended ==> the <b>[R programming language](https://en.wikipedia.org/wiki/R_)</b> and <b>[GitHub](https://en.wikipedia.org/wiki/GitHub)</b>.    

* Data should be downloaded from original sources, then analyzed, tabulated, mapped, and graphed via programs writen in R. Findings should be summarized in text laid out in a subset of <b>[Markdown](https://en.wikipedia.org/wiki/Markdown)</b> that an R application called <b>["knitr"](http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html)</b> converts to HTML pages suitable for publication on Websites. 

* Reproducible reports should be as concise as traditional reports, but the writer's notes, programs, analytical tools, coding guides, data files, tables, maps, graphs should be shared with the world via GitHub repositories. Of course Python, MATLAB, and other comparable languages could be used insted of R, and the supplementary files could be stored on other public repositories comparable to GitHub.     

This being my first attept to write a reproducible report, I wrote it in R and stored my supporting files in a public <b>[GitHub repo](https://github.com/Archimedes2kxx/BestStatesForBlackTech)</b>. If readers disagree with my numbers, they won't have to wonder how I got them <i>... #OldDogStillLearningNewTricks ... :-)</i>   

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>   

## Findings ... <i><small>Answers to the following questions</small></i>
My brief answers to the first three questions provide a broad statistical framwork for my extended responses to the fourth and fifth questions that form the core of this report.

#### Question 1A -- How many Whites, Blacks, Asians, and Hispanics lived in the U.S. in 2014?    
Answer 1 ... In 2014, as shown in Table 1A (below), the largest of the four racial/ethnic groups discussed in this report was White (`r prettyNum(dfTable1A[1,2], big.mark=",", scientific=FALSE)`); the smallest was Asian (`r prettyNum(dfTable1A[1,4], big.mark=",", scientific=FALSE)`). Hispanics, the largest minority (`r prettyNum(dfTable1A[1,5], big.mark=",", scientific=FALSE)`), had grown substantially larger than Blacks (`r prettyNum(dfTable1A[1,3], big.mark=",", scientific=FALSE)`).

<b>Table 1A -- Racial/Ethnic Composition of the U.S. Population in 2014</b>
```{r, echo=FALSE, cache=TRUE, comment=NA}
   print(format(dfTable1A, big.mark=","), row.names=TRUE)
```


#### Question 1B -- How many Whites, Blacks, Asians, Hispanics worked in tech in 2014? 
Answer 1B ... As shown in Table 2 (below), the U.S. Census Bureau estimated that there were `r prettyNum(dfTable1B[1,1], big.mark=",", scientific=FALSE)` tech professionals in the U.S. in 2014.  Whites were a substantial majority (`r prettyNum(dfTable1B[1,2], big.mark=",", scientific=FALSE)`) of the tech sector (`r prettyNum(dfTable1B[2,2], big.mark=",", scientific=FALSE)`%); Asians placed second (`r prettyNum(dfTable1B[1,4], big.mark=",", scientific=FALSE)`); Blacks showed third (`r prettyNum(dfTable1B[1,3], big.mark=",", scientific=FALSE)`); and Hispanics came in fourth  (`r prettyNum(dfTable1B[1,5], big.mark=",", scientific=FALSE)`)

<b>Table 1B -- Racial/Ethnic Composition of the U.S. Tech Sector in 2014</b>   
```{r, echo=FALSE, cache=FALSE, comment=NA}
  print(format(dfTable1B, big.mark=","), row.names=TRUE)
```

#### Question 2A -- What was the male/female composition of the U.S. population in 2014?

Answer 2A ... blah, blah, blah   

<b>Table 2A -- Male/Female Composition of the U.S. population in 2014</b>   
```{r, echo=FALSE, cache=FALSE, comment=NA}
  print(format(dfTable2A, big.mark=","), row.names=TRUE)
```

#### Question 2B -- What was the male/female composition of the tech sector in 2014? 

Answer 2B ... blah, blah, blah   

<b>Table 2B -- Male/Female Composition of the U.S. Tech Sector in 2014</b>   
```{r, echo=FALSE, cache=FALSE, comment=NA}
  print(format(dfTable2B, big.mark=","), row.names=TRUE)
```


<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>       

#### Question 3A -- Which standard occupation categories specified "information technology" for this report? 
Answer 3A ... For the purposes of this report, "information technology" was defined as the thirteen standard occupation categories (SOC) that appear in the first column of Table 3 (below). Eleven of these categories related to software, e.g., Software Developers, Computer Programmers, and Database Administrtors; only two related to hardware, i.e., Electrical and Electronics Engineers and Computer Hardware Engineers.  

#### Question 3B -- How many techs were in each category?        
Answer 3B ... The total number of techs in each category appear the second column of Table 3 (below). The top row shows that there were approximately `r prettyNum(dfTable3[1,2], big.mark=",", scientific=FALSE)` professionals in  all tech categories.

#### Question 3C -- What was the male/female breakdown of employment within these categories?         
Answer 3C ... The third column of Table 3 (below) shows the percentage of techs in the second column who were male. Females were an overall minority of `r dfTable3[1,6] `% of all professionals in tech, and were a minority in every tech category, ranging from a low value of `r prettyNum(min(dfTable3[, 6]), scientific=FALSE)`% for to a high value of `r prettyNum(max(dfTable3[, 6]), scientific=FALSE)`%.

<b>Table 3 -- Employment in U.S. Information Technology in 2014 with Female Percentages</b>   
```{r, echo=FALSE, cache=FALSE,  comment=NA}
   print(format(dfTable3[,c(1:2,6)], big.mark=","),row.names=FALSE) 
```
<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>    

#### Question 4 -- Where did the nation's Black, Hispanic, and female tech professionals work in 2014?      
Answer 4 ... The following sections provide an extensive response to a more specific formulation of this question ==> In which states did White, Black, Asian, Hispanic, and female tech professionals reside in 2014? <b>My underlying assumption was that most people worked in the states in which they lived. So the number of tech residents in a state should be about the same as the number of techs employed in that state.</b> (<i>Readers are reminded of two notable exceptions: New York City and Washington, DC wherein substantial numbers of out-of-state workers commute from nearby New Jersey and Connecticut, and from Maryand and Viginia, respectively.</i>) 

Table 4 (below) lists the six states that employed the most techs in 2014. These states will also appear in just about every table that follows because  they employed such a large share (`r prettyNum(min(dfTable4[1, 9]))` percent) of all `r prettyNum(dfTable4[1, 1], big.mark=",", scientific=FALSE)` techs in the U.S.

<b>Table 4 ... Biggest Half-Dozen Tech States in 2014</b>
```{r, echo=FALSE, cache=FALSE, comment=NA}
rownames(dfTable4) <- ""
### 1-row table is too long to print on one row, so totals in drop col 8
print(format(dfTable4[,c(1:7,9)], big.mark = "," , rownames=FALSE ))
```

I've presented U.S. Census data that shows the geographic distribution of tech professionals in two formats: (1) as a set of choropleth maps and (2) as a corresponding set of tables. I show the maps first because it's easier to see the residential patterns of the racial/ethnic/female groups by looking at multicolored maps than by scanning columns of data in long tables. Indeed, a quick scan of the maps makes it easier to see the same patterns in the data tables.

There are six maps (below), one for each racial/ethnic group plus two female groups, Asian females and non-Asian females. Map 4A shows the residential distribution of White techs; Map 4B shows the distribution of Black techs; Map 4C shows Asian techs; Map 4D shows Hispanics; Map E shows Asian females, and Map F shows non-Asian females. Each map colors the states with the largest number of tech residents in the darkest colors, and colors the states with the least tech residents in the lightest colors. The same colors represent the same percentages on all six maps

<b> Maps 4A (Whites), 4B (Blacks), 4C (Asians), 4D (Hispanics), 4E (Asian Females), and 4F (Non-Asian Females)</b> ... <em>Lower 48 and DC</em>
```{r, echo=FALSE, cache=FALSE, comment=NA}
    grid.arrange(dfMap4A, dfMap4B, dfMap4C, dfMap4D, dfMap4E, dfMap4F, ncol=2) 
```

* Map 4A shows that White techs are spread across a large number of states, with the largest number living in California, second highest in Texas. 

* Map 4B shows that most Black techs are located in the Eastern and Southern states, with the largest number living in Texas, second in Georgia, and third in Virginia

* Map 4C shows that Asian techs are highly concentrated in California where their concentration is higher any other group in any state. Otherwise Asians are spread out over other states with much smaller Asian residencies.   

* Map 4D shows that Hispanic techs are concentrated in California, Texas, and Florida   

* Map 4E shows that Asian female techs are highly concentrated in California; but substantial percentages also live in Texas and Florida.

* Map 4F shows that non-Asian females are have their largest shares of tech in California, Texas, and Florida, but less so than Asian females.

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>

<b>Tables 4A, 4B, 4C, 4D, 4E, and 4F</b> ... <i>50 states plus DC</i>  
Six data tables impose numeric precision on the qualitative insights provided by the corresponding maps. The tables are ordered by the number of techs from each group that lived in each state. The short versions that only contain the top ten rows are displayed below. To view the full tables that contain data for all 50 states plus the District of Columbia, please click the link below the table name. Each table (full and short version) has the following eight columns<br /><br />
1. <b>State</b> -- State of residence for the techs whose data is shown in the other columns of each row<br /> <br />
2. <b>AllTech</b> -- Data for all of the techs in the U.S. The data in this column is repeated in each of the four tables, but the order of the rows differs from one table to the next. For example Table 4A shows that the total number of techs in California in 2014 was `r prettyNum(dfTechPop_white["California", 2], big.mark=",", scientific=FALSE)`, and the total number in Texas was   `r prettyNum(dfTechPop_white["Texas", 2], big.mark=",", scientific=FALSE)` <br /> <br />
3. <b>S_%</b> -- The state's share of each racial/ethnic group's national total. More specifically, the percent of all U.S. techs in a racial group who lived in each state. For example, Table 4A shows that  `r dfTechPop_white["California", 3]`  percent of all White techs lived in California in 2014, and `r dfTechPop_white["Texas", 3]`  percent lived in Texas.<br /> <br />
4. <b>WhiTech, BlkTech, AsiTech, HispTech, FemAsiTech, or FemNonAsiTech</b> -- White techs, Black techs, Asian techs, Hispanic, femal Asian techs, and female Non-Asian techs in tables A, B, C, D, E, and F, respectively. The data in this column shows the number of techs from the table's racial/ethnic group that lived in each state. For example, Table 4A shows that `r prettyNum(dfTechPop_white["California", 4], big.mark=",", scientific=FALSE)` White techs lived in California, and `r prettyNum(dfTechPop_white["Texas", 4], big.mark=",", scientific=FALSE)` White techs lived in Texas. <b>The rows of each table are ordered by data in this column, i.e., by the number of techs in each state; states with higher numbers at the top, lower numbers at the bottom.</b><br /> <br />
5. <b>T_%</b> -- Percent of of the state's techs who were in the table's racial/ethnic group. For exampleTable 4A shows that `r dfTechPop_white["California", 5]`  percent of the techs in California were White, and `r dfTechPop_white["Texas", 5]`  percent of the techs in Texas were White.<br /> <br />
6. <b>WhiPop, BlkPop, AsiPop, HspPop, FemAsiTech, or FemNonAsiTech</b> -- White population, Black population, Asian population, Hispanic, female Asian population, or female non-Asian population in each state. For example Table 4A shows that the 2014 White (non-Hiispanic) population of California was `r prettyNum(dfTechPop_white["California", 6], big.mark=",", scientific=FALSE)`, and the White (non-Hispanic) population of Texas was `r prettyNum(dfTechPop_white["Texas", 6], big.mark=",", scientific=FALSE)`.<br /><br />
7. <b>P_%</b> -- Percent of the state's total population that belonged to the table's racial/ethnic group. For example, Table 4A shows that in 2014,  `r dfTechPop_white["California", 7]` percent of the total population of California was White, and `r dfTechPop_white["Texas", 7]` percent of the total population of Texas was white<br />.<br />
8. <b>Par</b> -- Parity ratio of the  percent of the state's tech population that belonged to the table's racial/ethnic group divided by the percent of the state's total population that belonged to that group, i.e., T_% / P_%. For example, Table 4A shows that the White parity ratio for California was  `r dfTechPop_white["California", "parity"]`, which is `r dfTechPop_white["California", 5]` divided by `r dfTechPop_white["California", 7]`<br /><br />

<b>Table 4A -- Whites in U.S. Information Technology -- Top 10 States </b>   
<i>[Click here for full version](Table4A-Full-White-Employment.html)</i>
```{r, echo=FALSE, cache=TRUE, comment=NA}
    colnames(dfTable4A) <- c("State", "AllTech", "S_%", "WhiTech", "T_%", "WhiPop", "P_%", "Par")
print(format(dfTable4A[1:11,], big.mark = ","), row.names=FALSE)  

```
<b>Comments about Table 4A -- Whites in Tech</b>   

* The numbers in the third (S_%) and fourth (WhiTech) columns of this table, i.e., the percentage of the nation's White techs who lived in each state and the estimated number of techs (WhiTech) in each state are equivalent to the colors on Map 4A. Indeed, the colors on this map are based on the numbers in these columns.   

* To be specific, the third and fourth columns show that in 2014 the largest number of White techs lived in California; the next largest lived in Texas. White techs were found in every state after that in smaller and smaller numbers.   

* This story is captured by the parity values in the last column (Par) of Table 4A (and in its full version). White parity values were all close to 1.0, which means that the percentage of techs who were White in each state was more or less the same as the percentage of Whites in the state's total population. In other words, if you walked into a tech office in any state in the USA, you would probably encounter the same percentage of White techs as the percentage of White people you would encounter walking down the street -- with an important exception. Random encounters with 10 White people on a street would only involve five males; but as per Table 3 (above) random encounters with 10 White techs in an office would probably involve seven or eight males.
<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>  

<b>Table 4B -- Blacks in U.S. Information Technology -- Top 10 states</b>    
<i>[Click here for full version](Table4B-Full-Black-Employment.html)</i> 
```{r, echo=FALSE, cache=TRUE, comment=NA}
    colnames(dfTable4B) <- c("State", "AllTech", "S_%", "BlkTech", "T_%", "BlkPop", "P_%", "Par")  
    print(format(dfTable4B[1:11,], big.mark = ","), row.names=FALSE)  
```
<b>Comments about Table 4B -- Blacks in Tech</b>   

* As I noted earlier, New York City and Washington, DC have large daytime commuter populations. Therefore the number of Black techs who worked in New York state or Washington, DC in 2014 may have been substantially larger than the number of Black techs listed in Table 4B.

* Column 4 (BlkTech) of Table 4B (above) shows that in 2014 the largest number of Black techs lived in Texas, second largest in Georgia, third in Virginia, fourth in New York, fifth in Maryland, sixth in Florida, and California placed seventh. Why? Because in 2014 the vast majority Black Americans still lived in Eastern and Southern states; so that's where most of the Black techs were found. This put most Black techs at a considerable disadvantage with regards to learning about and/or being recruited for tech jobs in Silicon Valley.

* California's prominence as the corporate headquarters of many of the nation's richest tech companies makes it necessary to closely examine the status of its disadvantaged groups. Unfortunately the numbers were disappointing. As per the fifth column (T\_%), the `r prettyNum(dfTechPop_black["California", "per_blackTech"], big.mark=",", scientific=FALSE)`%  Black share of tech in California was the lowest in the table by a wide margin. And its parity numbers in the last column (Par) were lower than all but one of the other nine states.    

* Perhaps California's `r prettyNum(dfTechPop_black["California", "blackPop"], big.mark=",", scientific=FALSE)` Black population wasn't large enough compared to the state's `r prettyNum( dfTechPop_black["California", "blackPop"]/dfTechPop_black["California", "per_blackPop"] * 100, big.mark=",", scientific=FALSE)` total population for Black techs to obtain a bigger slice of the tech pie? Unfortunately, this "critical mass" hypothesis is undermined by even lower proportions for Hispanics in Table D (below). By 2014 Hispanics had become the largest group in the state, `r prettyNum(dfTechPop_hispanic["California", "per_hispanicPop"], big.mark=",", scientific=FALSE)`%, larger than any plausible "critical mass". The Hispanic share of the state's total population was about seven times as large as the Black share, `r prettyNum(dfTechPop_black["California", "per_blackPop"], big.mark=",", scientific=FALSE)`%. So its tech share should have been at least seven times as large ... but it wasn't. As per Table 4D, the Hispanic tech share, `r prettyNum(dfTechPop_hispanic["California", "per_hispanicTech"], big.mark=",", scientific=FALSE)`%, was less than four times as large as the Black share, `r prettyNum(dfTechPop_black["California", "per_blackTech"], big.mark=",", scientific=FALSE)`%. So lack of "critical mass" can't explain either group's low share of California's tech sector. Other factors must have been involved.


<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small> 

<b>Table 4C -- Asians in U.S. Information Technology -- Top 10 states</b>    
<i>[Click here for full version](Table4C-Full-Asian-Employment.html)</i>

```{r, echo=FALSE, cache=TRUE, comment=NA}
    colnames(dfTable4C) <- c("State", "AllTech", "S_%",  "AsnTech", "T_%", "AsnPop", "P_%", "Par")
    print(format(dfTable4C[1:11,], big.mark = ","), row.names=FALSE)  
```
<b>Comments about Table 4C -- Asians in Tech</b>    

* As per the second row in the fourth column in Table 4C, California was home to `r prettyNum(dfTechPop_asian["California", "asianTech"], big.mark=",", scientific=FALSE)` Asian techs, which represented `r dfTechPop_asian["California", "per_state"]`x percent of all of the Asian techs in the U.S. This was, by far, the highest concentration of any racial/ethnic group in any state. Asian techs were also `r dfTechPop_asian["California", "per_asianTech"]`x percent of the `r prettyNum(dfTechPop_asian["California", "totalTech"], big.mark=",", scientific=FALSE)`x techs in California, the highest percentage of Asians in the tech sector of any state. (See the <i>[full version](Table4C-Full-Asian-Race-Residence.html)</i> of Tablc 4C)

* Although the intensity of Asian invovement in the tech sector of California was impressive, Asian participation in the tech sector of every other state was also remarkable. The parity values in the last column of the table provide concise summaries. The <i>[full version](Table4C-Full-Asian-Race-Residence.html)</i> of Table 4C shows that Asian parity values  ranged from `r dfTable6["Asian", "min"]` to `r dfTable6["Asian", "max"]`. In other words, the Asian share of the tech sectors in the 50 states plus DC ranged from `r dfTable6["Asian", "min"]` times their share of the state's total population to `r dfTable6["Asian", "max"]` times their share of the state's total population.   

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small> 

<b>Table 4D -- Hispanics in U.S. Information Technology -- Top 10 states</b>      
 <i>[Click here for full version](Table4D-Full-Hispanic-Employment.html)</i> 

```{r, echo=FALSE, cache=TRUE, comment=NA}
    colnames(dfTable4D) <- c("State", "AllTech", "S_%", "HspTech", "T_%", "HspPop", "P_%", "Par")
    print(format(dfTable4D[1:11,], big.mark=","), row.names=FALSE)  
```
<b>Comments about Table 4D -- Hispanics in Tech</b>  

* As per the third column (S_%) of Table 4D, Hispanic techs were concentrated in three states: California, Texas, and Florida. The table shows that `r dfTechPop_hispanic["California", "per_state"]`x percent lived in California,  `r dfTechPop_hispanic["Texas", "per_state"]`x percent lived in Texas, and  `r dfTechPop_hispanic["Florida", "per_state"]`x percent lived in Florida. Taken together, these three states accounted for `r dfTechPop_hispanic["California", "per_state"] + dfTechPop_hispanic["Texas", "per_state"] + dfTechPop_hispanic["Florida", "per_state"] `x percent of the Hispanic techs in the entire country, i.e., about half. The other half live in the other 47 states and the District of Columbia.  

* All of the parity values of the top 10 states for Black techs shown in the last column of Table 4B are substantially higher than the parity values for all of the top 10 states for Hispanic techs shown in Table 4D ... with one exception: Florida, parity = `r dfTechPop_hispanic["Florida", "parity"]`x. Florida's parity value for Hispanic techs was more than twice the `r dfTechPop_hispanic["California", "parity"]`x parity value for Hispanics in California. Indeed, Table 4D shows that California had the lowest parity value of all of the top ten states for Hispanics. These considerations suggests that Hispanic techs may have encountered larger employment barriers than Black techs, especially in California.   

* Despite California's relatively low Hispanic parity, the fifth column (T_%) of Table 4D shows that `r dfTechPop_hispanic["California", "per_hispanicTech"]`x percent of the techs who lived in California were Hispanic. This figure is remarkable because it is two to three times as high as the 2 percent to 4 percent employment rates for Hispanic techs that have been reported by Silicon Valley's biggest firms in the last three years. It suggests that willingness to hire Hispanic techs was low throughout the state, but substantantially lower in the Valley.    

* Perhaps the Valley required higher academic qualifications than other employers. But as I pointed out in a recent note on this blog (<i>["The California Pipelines of Silicon Valley"](http://hbcu-levers.blogspot.com/2016/07/the-california-pipelines-of-silicon.html)</i>), IPEDS data showed that 20 percent of the combined total of White, Black, Asian, and Hispanic students enrolled in the best colleges and universities in California were Hispanic. So there is reason to be skeptical about claims that few Hispanic techs met the Valley's higher recruiting standards. 


<b>Table 4E -- Asian Females in U.S. Information Technology -- Top 10 states</b>     
 <i>[Click here for full version](Table4E-Full-AsianFemale-Employment.html)</i> 

```{r, echo=FALSE, cache=TRUE, comment=NA}
    colnames(dfTable4D) <- c("State", "AllTech", "S_%", "AsiFemTech", "T_%", "NonAsiFemPop", "P_%", "Par")
    print(format(dfTable4D[1:11,], big.mark=","), row.names=FALSE)  
```
<b>Comments about Table 4E -- Asian Females in Tech</b>    
 
 * blah blah
 * blah blah


<b>Table 4F -- Non-Asian Females in U.S. Information Technology -- Top 10 states</b>    
 <i>[Click here for full version](Table4F-Full-NonAsianFemale-Employment.html)</i> 

```{r, echo=FALSE, cache=TRUE, comment=NA}
    colnames(dfTable4F) <- c("State", "AllTech", "S_%", "NonAsiTech", "T_%", "NonAsiPop", "P_%", "Par")
    print(format(dfTable4F[1:11,], big.mark=","), row.names=FALSE)  
```
<b>Comments about Table 4F -- Non-Asian Females in Tech</b>  
 
 * blah blah
 * blah blah

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small> 

### Question 5 -- Was the size of a group's share of a state's population related to the size of the group's share of the tech jobs in that state? 

Answer 5: As shown in four plots that follow, there was a positive relationship, but the relationship was strongest for Asians, weaker for Asian females, still weaker for Whites, weaker for Blacks, weaker for non-Asian females, and weakest for Hispanics

<b>Plots 5 -- Plots of Tech Employment vs Race/Ethnic/Female Group Population</b>   
```{r, echo=FALSE, cache=FALSE, warning=FALSE}
    grid.arrange(ggPlot_white, ggPlot_black, ggPlot_asian, ggPlot_hispanic, ggPlot_femAsian, ggPlot_femNonAsian, ncol=2)
```

<b>Comments about Plots 5A, 5B, 5C, 5D, 5E, 5F</b> 

* The total population of each state was at least 1,000 times the population of the techs in any racial/ethnic group. Therefore I ran simple regressions of techs in a group against the total population of that group divided by 1000 in all 50 states plus Washington, DC. The slopes of the regression lines provided estimates of the number of additional techs that would be found in a state if the size of the group increased by 1000. 

* The regression lines and scatterplots of the <tech, population> data points appear in the four panels. The values of the slopes, called "beta1000", are displayed in the upper left corner of each plot. 

* As per the values of the beta1000 slopes, If each ethnic/racial group's population increased by 1,000, Asian techs would increase by `r beta1000["asian"]`; Whites would increase by `r beta1000["white"]`, Blacks would increase by `r beta1000["black"]`; and Hispanics would increase by `r beta1000["hispanic"]`. The Asian line has the steepest slope; the Hispanic line has the flattest slope. 

* This kind of simple regression implicitly assumes that the 50 states and DC were all more or less equally receptive or resistant to including more techs from the same groups. This may indeed have been true for most states. However, Table 4B shows that the Black parity for Texas = `r dfTechPop_black["Texas", "parity"]`, whereas the Black parity for California = `r dfTechPop_black["California", "parity"]`. This suggests that Texas was more receptive to Black techs than California. And Table 4D shows that the Hispanic parity for Florida = `r dfTechPop_hispanic["Florida", "parity"]`, while the Hispanic parity  for California = `r dfTechPop_hispanic["California", "parity"]`. This suggests that Florida was more receptive to Hispanic techs than California.   

* blah, blah, blah about 5E and 5F

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small> 


<b>Table 5 -- Summary of Racial/Ethnic/Female National Advantages</b>
```{r, echo=FALSE, cache=FALSE, comment=NA}
    print(format(dfTable5, big.mark = ","), row.names=TRUE) 
```

<b>Comments about Table 5</b>

* The parity and beta1000 statistics shown in Table 5 provide different ways to measure the same concept ==> The relative advantages in the sector that the nation as a whole provided to the four racial/ethnic groups in 2014. 

* The percentages and ratios shown in Table 5 are national percentages and national parity ratios -- not the state percentages and state parity ratios shown in Tables 4A, 4B, 4C, 4D, 4E, and 4F. In other words column 2 (T_%) shows the percentage of all techs in the nation who were in each group. Column 4 shows the percentage of the nation's entire population that was in each group. The national parity ratios for each group shown in column 5 (Par) are the result of dividing the national tech percentages in column 2 (T_%) by the national population percentages in column 4 (S_%). 

* As per the plots shown in this section, the more complex beta1000 statistics were the results of regressions that estimated a linear relationship between the size of a group's tech population (T_% in Tables A, B, C, D, E, and F) in every state to the group's total population in every state  (S_% in Tables A, B, C, D, E, and F).    

* As can be seen from the last two columns in the table, the two measures provided similar results: Asians were the most advantaged {parity = `r dfTable5["Asian", "Par"]`, beta1000 = `r dfTable5["Asian", "beta1000"]` }, Whites second {parity = `r dfTable5["White", "Par"]`, beta1000 = `r dfTable5["White", "beta1000"]`}, Blacks third {parity = `r dfTable5["Black", "Par"]`, beta100 = `r dfTable5["Black", "beta1000"]`}, and Hispanics were the least advantaged {parity = `r dfTable5["Hispanic", "Par"]`, beta100 = `r dfTable5["Hispanic", "beta1000"]`}. Asian females {parity = `r dfTable5["FemAsian", "Par"]`} were more advantaged than Whites {parity = `r dfTable5["White", "Par"]`}, and far more advantaged than non-Asian females {parity = `r dfTable5["FemNonAsian", "Par"]`}

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>

## Conclusions    
Which states were the best states for Black, Hispanic, and female professionals in information technology? Given the limited scope of the data that I examined in this study, my selection criteria were rudimentary; so my conclusions should be regarded as tentative first approximations. Rather than just address this question for the disadvantaged Black, Hispanic, and female techs, I answered it for all six groups because the answers for the other three groups provided context for my conconclusions about the disadvantaged groups.

1. <b>Select states that had the largest numbers of techs in the racial/ethnic groups and the largest numbers Asian females and non-Asian females</b>   
I implemented this criterion by selecting the states with the ten highest numbers of techs in each group. This meant that I began with the states that appeared in short versions of Tables 4A through 4F.

2. <b>Select states that had the largest parity values</b>   
I implemented this criterion by identifying the states in Tables 4A through 4F that had state level parity values in the top 50% of all 50 states plus DC. Table 6 (below) summarizes the quartile breakdowns of the parity values for each group. For example, Table 6 shows the median Black parity value was  `r dfTable6["Black", "median"]`;  the median Hispanic parity value was `r dfTable6["Hispanic", "median"]`.   

<b>Table 6 -- Summary of Parity Values</b>
```{r, echo=FALSE, cache=FALSE, comment=NA}
    colnames(dfTable6) <- c("Min", "Q1", "Median", "Q3", "Max")
    print(dfTable6)  
```
3. <b>Identify the finalists</b>   
Table 7A through 7F lists the states whose parity values were greater than or equal to the median parity value for each group. 

<b>Tables 7A, 7B, 7C, 7D, 7E, 7F -- Finalists</b>
```{r, echo=FALSE, cache=FALSE, comment=NA}
    ###print(dfTable7F)  
    grid.arrange(tableGrob(dfTable7A), tableGrob(dfTable7B), tableGrob(dfTable7C), tableGrob(dfTable7D), tableGrob(dfTable7E), tableGrob(dfTable7F), nrow=3) 

```


4. <b>Best States</b>    
I chose the "Best States" by sorting the lists of finalists in decreasing order of each group's percentage share (T_%) of the tech sector. My original intention was to select the top five finalists according to this ranking. But no list had more than five finalists; so all finalists were selected for each group. The state at the top of each list was the best of the best for its group, the state in the second row was the second best, and so on. Readers should remember that <b>my underlying assumption was that most people worked in the states in which they lived. So the number of tech residents in a state should be about the same as the number of tech jobs in a state.</b>. 

<b>Tables 8A, 8B, 8C, 8D, 8E, 8F -- Best States</b>
```{r, echo=FALSE, cache=FALSE, comment=NA}
        grid.arrange(tableGrob(dfTable8A, theme=tt3), tableGrob(dfTable8B, theme=tt3), tableGrob(dfTable8C, theme=tt3), tableGrob(dfTable8D, theme=tt3), tableGrob(dfTable8E, theme=tt3), tableGrob(dfTable8F, theme=tt3), nrow=3) 
```


## Comments about Version 0.9 and additional content for Version 1.0</b>   

* Lies, Damned Lies, and Statistical Categories

... Strength of weak ties, old tech states don't make anyone's best lists except Whites ... old White Boy network

* Version 1.0 -- H1-B visas for Asians 
.... Palantir



<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>   

## Methods   
I downloaded the  <b>[#PEPSR6H](http://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=PEP_2015_PEPSR6H&prodType=table)</b> data table from the Bureau's American Fact Finder Website. This table could be configured to provide the Bureau's estimates of the overall racial composition of the populations of the 50 states plus Washington DC in 2014. My step-by-step procedure for downloading the AFF data is described in <b>[Notes-Data-Raw-Maps-Plots](https://archimedes2kxx.github.io/BestStatesForBlackTech/Notes-Data-Raw-Maps-Plots)</b> on the GitHub repo.

American Fact Finder contained a wide range of tables, but none of them broke the information technology sector down by race, ethnicity, and state. So I also used the U.S. Census Bureau's <b>[DataFerret](http://thedataweb.rm.census.gov/index.html)</b> Java
application to download a subset of the Public Use Microdata Sample (PUMS) from the Bureau's American Community Survey for 2014. The subset (`r prettyNum(dim(dfCensus2)[1], big.mark=",", scientific=FALSE)` observations) was usually large enough to make good estimates of the employee counts and percentages of tech employment for Blacks, Whites, Asians, and Hispanics in tech occupations in the 50 states and DC. (<i>Note: For a few low population states, the ACS subset included missing values which meant that it wasn't large to provide reliable estimates of the racial/ethnic breakdown of tech employment in those states</i>.)  My step-by-step procedures for selecting and downloading the ACS data and its codebook is described in <b>[Notes-Data-Download-PUMS-Sample](https://archimedes2kxx.github.io/BestStatesForBlackTech/Notes-Data-Download-PUMS-Sample.html)</b> on the GitHub repo.<br />

<b>Selected Data</b><br/> 
I used the DataFerret to download a subset of the PUMS 2014 sample about the sex, state of residence, race, and Hispanic status of employees in the tech sector in all 50 states plus the District of Columbia:<br />   

* <b>Occupation</b><br />When using the DataFerret app to select a sample of tech professionals from the full PUMS 2014 sample, I defined the "information technology sector" by designating the 13 standard occupational categories shown in Table 3. Eleven categories were software, two were hardware   

* <b>Hispanic vs White (not Hispanic), Black (not Hispanic), Asian (not Hispanic)</b><br />
I configured the DataFerret to include all races in my sample. However, Hispanic is not regarded as a race, but as an ethnic/national origin, like being Irish, Egyptian, Japanese, or Nigerian. So I included the "Hispanic" variable. This variable designated each respondent's Hispanic origin, e.g., Mexico, Chile, Puerto Rico, etc ... but it also contained a "Not Hispanic" value for respondents who were not from Hispanic countries. Using this code, I distinguished between "White" vs. "White (not Hispanic)", "Black" vs. "Black (not Hispanic)", and "Asian" vs. "Asian (not Hispanic)" and "Others" vs "Others (non-Hispanic)". Accordingly, my report contains data about four racial/ethnic groups: Hispanic, White (not Hispanic), Black (not Hispanic), and Asian (not Hispanic); but for brevity it merely calls them Hispanic, White, Black, and Asian.

* <b>Estimating population values from the sample</b>   
Each row in the downloaded PUMS table contained the anonymized responses of a real tech employee in the ACS 2014 survey, i.e., sex, race/ethnicity, occupation, and state of residence. Each observation also contained a number called "personal weight" which is the Census Bureau's estimate of how many people in the real population were like each person in the sample. An estimate of the total number of techs in the real population was found by adding up the personal weights of the techs in the sample. This also applied to subgroups, e.g, male vs. female, state of residencere, etc. The percentage share of a subgroup's residence was obtained by dividing the estimated residence in the subgroup by the estimated residence in the entire group, i.e., dividing the sum of personal weights in a subgroup by the sum of personal weights in the larger group. <br /><br />
Here's the <b>[Census Bureau's description](https://usa.ipums.org/usa/voliii/ACSsamp.shtml)</b> of this process:<br /><i>"Production of Estimates"</i> -- The ACS PUMS sample is not self - weighted. To produce estimates or tabulations of characteristics from the ACS PUMS simply add the weights of all persons or housing units that possess the characteristic of interest. For instance, if the characteristic of interest is "total number of black teachers", simply determine the race and occupation of all persons and cumulate the weights of those who match the characteristics of interest. To get estimates of proportions simply divide the weighted estimate of persons or housing units with a given characteristic by the weighted estimate of the base. For example, the proportion of "black teachers" is obtained by dividing the weighted estimate of black teachers by the PUMS estimate of teachers."
   
I used the <b>[R Studio IDE](https://www.rstudio.com/)</b> to write the R scripts and functions that produced the report's tables, maps, graphs, and stats. My R code can be found in the Data-1A.R, Data-1B.R, Stats-2.R, and Report-3.Rmd files in the <b>[Best States ](https://github.com/Archimedes2kxx/BestStatesForBlackTech)</b> repo, as well as copies of the codebooks and .csv files I downloaded from the Census. I composed the text of the report in R-Markdown, embedded the data objects, converted the R-markdown to HTML via knitr, pushed the report to git-io pages, then made the pages accessible to readers of the DLL's <b>[Tech-Levers](http://hbcu-levers.blogspot.com/p/draft-of-best-states-for-blacks-in-tech.html)</b> blog via an iframe on a page on the blog. 

The scripts and embedded functions should be run in numerical order ==> Data-1A.R and Data-1B.R create the data frames required by Stats-2.R; Stats-2.R creates the tables, maps, graphs, and regression stats required by the R-markdown script in the Report-3.Rmd file that generates the report.

Finally, this report represents my honest attempt to shed additional light on some important, but difficult issues. Therefore I would greatly appreciate readers' comments and suggested improvements ... but I would cherish their help in identifying any errors that I made in my data collection or analysis because bad reports don't support good policies ... :-)
<br />
<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small> 

