---
title: "Foreign Professionals: U.S. Tech's Other Pipelines -- Version 0.4"
author: "Roy L Beasley, PhD -- DLL Editor"
date: "12/19/2016"

output: 
 prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
    toc_depth: 4
    fig_width: 7
    fig_height: 6
                  
---

```{r LOAD, echo=FALSE, cache=FALSE}

    #### load("dfRaceWorkplaceAndShares.rda") ### Only use data in tables, plots, and maps
    load("dfCensus2.rda") 
    load("dfTab1A1B2A2B3ABCDEF.rda")
    load("dfTab4.rda")   
    load("dfMap4.rda")
    load("dfPlot5Tab5beta1000.rda")
    load("dfTab67A7B7C7D7E7F8A8B8C8D8E8F.rda")
    load("APPENDIX.rda")  
    
    ###install.packages("gridExtra")
    library(prettydoc)
    library(maps)    
    library(mapproj) ### needed by ggplot2, but not installed automaticallly
    library(ggplot2)
    library(gridExtra)
    ### library(printr)
    library(grid)
    library(gtable)
    library(grid)

    ### Intending to use Yihui Xie's printr package for better tables later
    ### install.packages("devtools")
    ### library(devtools)
    ### install_github("yihui/printr")
    ### library(printr)
    ### detach('package:printr', unload = TRUE)
    
tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = blues9[1:7], col=NA),
            fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="navyblue", fontface=4L)),
  rowhead=list(fg_params=list(col="orange", fontface=3L)))

fem10 <- round(dfTable3ZZ[1, "Tech10"] * dfTable3ZZ[1, "perF10"]/100, digits=0) ### calculate females in 2010 from female share
perFemIncrease <- round(100 * (dfTable3Z[1,"Fem"] - fem10)/fem10, digits=1)

```
\pagenumbering{arabic}   

## Overview     
<i>Please note that all of the numbers in this overview and in the subsequent sections are estimates based on U.S. Census survey samples. So all numbers should be read as "approximately" even when the word "approximately" does not appear. </i>  

* blah, blah, blah   

* See Overview of full report <a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html#overview" target="_self"><b>HERE</b></a>


<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small>   
  
<P style="page-break-before: always">  
## Findings   

Brief answers to the first three sets of questions provide a broad statistical framwork for the extended responses to the fourth question that forms the core of this supplement.   

<i>Please note that the tables, maps, and charts in this supplement about foreign techs have been assigned numbers that follow the numbers of similar tables, maps, and charts in the original report about American techs. For example, Tables 3E and 3F appear in ths supplement because Tables 3A, 3B, 3C, and 3D appeared in the original. On the other hand a few tables from the original are repeated in this supplement as convenient references for the reader and are assigned the same numbers as in the original; for example, Tables 1C and 1E.</i>


#### Question 1 -- How many foreign professionals worked in the U.S. tech sector in 2015?

<i>Answer 1C</i> ...As can be seen from Table 1C (below), `r dfTable1C["Per", "Foreign"]` percent of the U.S. tech sector in 2015 were foreign professionals.  

<b>Table 1C -- American and Foreign Components of the U.S. Tech Sector in 2015</b>    
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(dfTable1C, row.names=TRUE, quote=FALSE)
```

<i>Question 1E</i> -- How many Asian and non-Asian foreign employees were in U.S. tech in the 50 states plus the District of Columbia in 2015? 

<i>Answer 1E</i> ...  As per Table 1E (below), there were almost three times as many foreign professionals from Asian countries as from other countries in U.S. tech in 2015.
  
<b>Table 1E -- Foreign Asian and Non-Asian Components of the U.S. Tech Sector in 2015</b> 
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(dfTable1E, row.names=TRUE, quote = FALSE)
```
<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small>   

<P style="page-break-before: always">   
#### Question 2 -- What was the foreign male/female composition of U.S. tech in 2015?
<i>Answer 2</i> ...Asper Table 2C ...
Table 2C ... tot male female femAsian femNonAsian

blah-blah-blah

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small>       

<P style="page-break-before: always">    
#### Question 3 -- How did the foreign shares of jobs in U.S. tech change between 2010 and 2015?    


The six pairs of tables in this section display profiles of the various groups considered in this report. 

For the purposes of this report, "information technology" was defined as the thirteen standard occupation categories (SOC) that appear in the first column of the following tables. Twelve related to software, e.g., Software Developers and Computer Programmers; one related to hardware, i.e., Computer Hardware Engineers.     

The first column of each table contains a list of job categories. Some categories were abbreviated to enable the rows of the table to be short enough to appear on the same line. Abbreviations include "Comp" for "Computer"; "Info" for "Information"; and "Net" for "Network". Column headings were also abbreviated as follows:   

* <b>Tech10</b> ... the number of tech workers in each job category in 2010   

* <b>Tech15</b> ... the number of tech workers in each job category in 2015   

* <b>perTS</b> ... the "Percentage of Tech Share", which shows the percentage of the total number of American tech workers that were employed in each job category in the 50 states plus the District of Columbia. For example, in Table 3Z (below) the `r prettyNum(dfTable3Z["SOFTWARE DEVELOPERS", "Tech15"], big.mark=",", scientific=FALSE)` workers in "Software Development" represented `r prettyNum(dfTable3Z["SOFTWARE DEVELOPERS", "perTS"], big.mark=",", scientific=FALSE)` percent of the `r prettyNum(dfTable3Z["All Occupations", "Tech15"], big.mark=",", scientific=FALSE)` total American tech workers in 2015. The `r prettyNum(dfTable3Z["COMP PROGRAMMERS", "Tech15"], big.mark=",", scientific=FALSE)` workers "Computer Programmers" were `r prettyNum(dfTable3Z["COMP PROGRAMMERS", "perTS"], big.mark=",", scientific=FALSE)` percent of the total `r prettyNum(dfTable3Z["All Occupations", "Tech15"], big.mark=",", scientific=FALSE)`.   

* <b>Fem</b> ... the number of "Female" tech workers". The number of males can be calculated by substracting the female workers from the total number of workers in each category. For example, in Table 3Z (below) there were `r prettyNum(dfTable3Z["SOFTWARE DEVELOPERS", "Tech15"], big.mark=",", scientific=FALSE)` American software developers in the 50 states plus DC in 2015 of which `r prettyNum(dfTable3Z["SOFTWARE DEVELOPERS", "Fem"], big.mark=",", scientific=FALSE)` were female. So the number of male software developers was `r prettyNum(dfTable3Z["SOFTWARE DEVELOPERS", "Tech15"], big.mark=",", scientific=FALSE)` - `r prettyNum(dfTable3Z["SOFTWARE DEVELOPERS", "Fem"], big.mark=",", scientific=FALSE)` = `r prettyNum((dfTable3Z["SOFTWARE DEVELOPERS", "Tech15"] - dfTable3Z["SOFTWARE DEVELOPERS", "Fem"]), big.mark=",", scientific=FALSE)`. 

* <b>perF15</b> ... the "Percent Female in 2015" displays the percent of jobs in each category that were held by female techs in 2015. For example, Table 3Z shows that the `r prettyNum(dfTable3Z["SOFTWARE DEVELOPERS", "Fem"], big.mark=",", scientific=FALSE)` female software developers were `r prettyNum(dfTable3Z["SOFTWARE DEVELOPERS", "perF15"], big.mark=",", scientific=FALSE)` percent of the total `r prettyNum(dfTable3Z["SOFTWARE DEVELOPERS", "Tech15"], big.mark=",", scientific=FALSE)`

* <b>perF10</b> ... the "Percent Female in 2010" displays the percent of jobs in each category that were held by female techs in 2010. This stat is always contained in the companion table, e.g., Table 3ZZ.

* <b>Change</b> ... the difference between the number of workers in each job category in 2010 and the number in 2015. This stat is always found in the companion table, e.g., Table 3ZZ. 

* <b>perCh</b> ... the change from 2010 to 2015 as a percentage of the jobs in 2010.  

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small>  

<P style="page-break-before: always">   
 
 <b>Table 3Z -- American Male/Female Composition of U.S. Tech in 2015</b> 
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(format(dfTable3Z, big.mark=","), row.names=FALSE, quote=FALSE)
```

<P style="page-break-before: always">    
 <b>Table 3ZZ -- Changes in the American Composition of U.S. Tech Between 2010 and 2015</b> 
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(format(dfTable3ZZ, big.mark=","), row.names=FALSE, quote=FALSE)
```
<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>  



<P style="page-break-before: always">   
 
 <b>Table 3E -- Foreign Male/Female Composition of U.S. Tech in 2015</b> 
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(format(dfTable3E, big.mark=","), row.names=FALSE, quote=FALSE)
```

<P style="page-break-before: always">    
 <b>Table 3EE -- Changes in the Foreign Composition of U.S. Tech Between 2010 and 2015</b> 
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(format(dfTable3EE, big.mark=","), row.names=FALSE, quote=FALSE)
```
<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>  



<P style="page-break-before: always">   
 <b>Table 3F -- Foreign Asian Male/Female Composition of U.S. Tech in 2015</b> 
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(format(dfTable3F, big.mark=","), row.names=FALSE, quote=FALSE)
```

<P style="page-break-before: always">    
 <b>Table 3FF -- Changes in the Foreign Asian Composition of U.S. Tech Between 2010 and 2015</b> 
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(format(dfTable3FF, big.mark=","), row.names=FALSE, quote=FALSE)
```
<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>  


<P style="page-break-before: always">   
 <b>Table 3G -- Foreign Male/Female Composition of Tech in California in 2015</b> 
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(format(dfTable3G, big.mark=","), row.names=FALSE, quote=FALSE)
```

<P style="page-break-before: always">    
 <b>Table 3GG -- Changes in the Foreign Composition of Tech In California between 2010 and 2015</b> 
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(format(dfTable3GG, big.mark=","), row.names=FALSE, quote=FALSE)
```
<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>  


<P style="page-break-before: always">   
 <b>Table 3H -- Foreign Asian Male/Female Composition of Tech in California in 2015</b> 
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(format(dfTable3H, big.mark=","), row.names=FALSE, quote=FALSE)
```

<P style="page-break-before: always">    
 <b>Table 3HH -- Changes in the Foreign Asian Composition of Tech In California between 2010 and 2015</b> 
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(format(dfTable3HH, big.mark=","), row.names=FALSE, quote=FALSE)
```
<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html" target="_self">back-to-top</a></center></i></small>     



Big 6 states ... 

Chart 5. Foreign Growth in Big Six States

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small>   

<P style="page-break-before: always">    
#### Question 4 -- Where did foreign techs work in 2015?      
Answer 4 ... The following sections provide an extensive response to a more specific formulation of this question ==> In which states did foreign tech professionals work in 2015?  
Map 4G Foreign   
Map 4H Asian   

Table 4G Foreign   
Table 4H Asian

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small>

<P style="page-break-before: always">  
## Final Comments</b>   

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small> 
    
<P style="page-break-before: always">  
## Personal Motivation   

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small> 
      
<P style="page-break-before: always">  
## Methods   

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small> 
 
<P style="page-break-before: always">    
## Appendix -- Profiles of Foreign Professionals in the Six Biggest Tech States in 2015

As per Table 4 in the supplement, the six states in this appendix employed the largest number of Americans in their tech sectors -- California, Texas, New York, Florida, Virginia, and Illinois. These six tables display the number of foreign techs employed by the Big Six.

Two pairs of tables are included for each state:     

* The first two tables show changes in the employement of foreign techs between 2010 and 2015 in the state's tech sector.   

* The second pair of tables show changes in employment of foreign Asian techs between 2010 and 2015 in the state's tech sector.   


In a future version of the supplement, this Appendix will be replaced by an interactive dashboard that will allow users to display pairs of tables for any state and for any American or foreign group.   

<i>Note: The column headings and abbreviated row names of the tables in this appendix are explained in the answer to [<b>Question 3](https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html#question-3-how-did-the-racial-ethnic-foreign-and-malefemale-shares-of-the-job-categories-in-u.s.-tech-change-between-2010-and-2015)</b> in the body of the supplement</i>



<P style="page-break-before: always">    
  
Table 3CA 3CACA   

Table 3TX 3TXTX    

Table 3NY 3NYNY    

Table 3FL 3FLFL    

Table 3VA 3VAVA      

Table 3IL 3ILIL    

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small> 

