---
title: "Foreign Professionals in U.S. Tech ... 'Other' Pipelines -- Version 0.2"
author: "Roy L Beasley, PhD -- DLL Editor"
date: "12/17/2016"

output: 
 prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
    toc_depth: 4
    fig_width: 7
    fig_height: 6
                  
---

```{r LOAD, echo=FALSE, cache=FALSE}

    #### load("dfRaceWorkplaceAndShares.rda") ### Only use data in tables, plots, and maps
    load("dfCensus2.rda") 
    load("dfTab1A1B2A2B3ABCDEF.rda")
    load("dfTab4.rda")   
    load("dfMap4.rda")
    load("dfPlot5Tab5beta1000.rda")
    load("dfTab67A7B7C7D7E7F8A8B8C8D8E8F.rda")
    load("APPENDIX.rda")  
    
    ###install.packages("gridExtra")
    library(prettydoc)
    library(maps)    
    library(mapproj) ### needed by ggplot2, but not installed automaticallly
    library(ggplot2)
    library(gridExtra)
    ### library(printr)
    library(grid)
    library(gtable)
    library(grid)

    ### Intending to use Yihui Xie's printr package for better tables later
    ### install.packages("devtools")
    ### library(devtools)
    ### install_github("yihui/printr")
    ### library(printr)
    ### detach('package:printr', unload = TRUE)
    
tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = blues9[1:7], col=NA),
            fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="navyblue", fontface=4L)),
  rowhead=list(fg_params=list(col="orange", fontface=3L)))

fem10 <- round(dfTable3ZZ[1, "Tech10"] * dfTable3ZZ[1, "perF10"]/100, digits=0) ### calculate females in 2010 from female share
perFemIncrease <- round(100 * (dfTable3Z[1,"Fem"] - fem10)/fem10, digits=1)

```
\pagenumbering{arabic}   

## Overview     
<i>Please note that all of the numbers in this overview and in the subsequent sections are estimates based on U.S. Census survey samples. So all numbers should be read as "approximately" even when the word "approximately" does not appear. </i>  

* blah, blah, blah   

* See Overview of full report <a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Report-3.html#overview" target="_self"><b>HERE</b></a>


<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small>   
  
<P style="page-break-before: always">  
## Findings   

Brief answers to the first three sets of questions provide a broad statistical framwork for the extended responses to the fourth question that forms the core of this supplement.   

<i>Please note that the tables, maps, and charts in this supplement about foreign techs have been assigned numbers that follow the numbers of similar tables, maps, and charts in the original report about American techs. For example, Tables 3E and 3F appear in ths supplement because Tables 3A, 3B, 3C, and 3D appeared in the original. On the other hand a few tables from the original are repeated in this supplement as convenient references for the reader and are assigned the same numbers as in the original; for example, Tables 1C and 1E.</i>


#### Question 1 -- How many foreign professionals worked in the U.S. tech sector in 2015?

<i>Answer 1C</i> ...As can be seen from Table 1C (below), `r dfTable1C["Per", "Foreign"]` percent of the U.S. tech sector in 2015 were foreign professionals.  

<b>Table 1C -- American and Foreign Components of the U.S. Tech Sector in 2015</b>    
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(dfTable1C, row.names=TRUE, quote=FALSE)
```

<i>Question 1E</i> -- How many Asian and non-Asian foreign employees were in U.S. tech in the 50 states plus the District of Columbia in 2015? 

<i>Answer 1E</i> ...  As per Table 1E (below), there were almost three times as many foreign professionals from Asian countries as from other countries in U.S. tech in 2015.
  
<b>Table 1E -- Foreign Asian and Non-Asian Components of the U.S. Tech Sector in 2015</b> 
```{r, echo=FALSE, cache=FALSE, comment=NA}
   print(dfTable1E, row.names=TRUE, quote = FALSE)
```
<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small>   

<P style="page-break-before: always">   
#### Question 2 -- What was the foreign male/female composition of U.S. tech in 2015?

<i>Question 2A</i> -- What was the foreign male/female composition of the tech sector in the 50 states plus the District of Columbia in 2015?

blah-blah-blah

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small>       

<P style="page-break-before: always">    
#### Question 3 -- How did the foreign shares of jobs in U.S. tech change between 2010 and 2015?     


<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small>   

<P style="page-break-before: always">    
#### Question 4 -- Where did foreign tech professionals work in 2015?      
Answer 4 ... The following sections provide an extensive response to a more specific formulation of this question ==> In which states did foreign tech professionals work in 2015?  


<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small>
    
<P style="page-break-before: always">  
## <i>"And the best states are ... "</i>


<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small> 
    
<P style="page-break-before: always">  
## Final Comments</b>   

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small> 
    
<P style="page-break-before: always">  
## Personal Motivation   

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small> 
      
<P style="page-break-before: always">  
## Methods   

<small><i><center><a href="https://archimedes2kxx.github.io/BestStatesForBlackTech/Supplement-3.html" target="_self">back-to-top</a></center></i></small> 
 
